\Title

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : B

To transform each movie into events, as recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames. For a given pixel and time stamp, an event is generated when the absolute value of this gradient exceeds a threshold. The event has either an OFF or ON polarity, depending on whether the gradient is negative or positive. The signed threshold is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, i.e. a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}). The goal here is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
%----------------------------%
%
\begin{figure}%[t!]
    \centering
    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons. Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}. When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function. In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}. The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron. An output spike is emitted when the membrane potential reaches the threshold and it is then reset.   \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses). }%
    \label{fig:izhikevich}%
\end{figure}%
% 
%----------------------------%
In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}. This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration). Events are typically ordered by their time of occurrence. Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$. This defines an address space $\presynaddrspace$ which consists of the set of possible addresses. In a neurobiological recording, this can be the identified set of presynaptic neurons. For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels. As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.

In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent neurons from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses. Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. Note that a neuron may contact another afferent neuron with different delays through different synaptic connections. Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses, 
as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$. This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of one neuron with synaptic weights and delays).

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. It results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$: 
%
\begin{linenomath*}
\begin{equation}\label{eq:stream_b}
\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
\end{equation}
\end{linenomath*}
%
In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons. However, it should be properly reordered in simulations. Crucially, when postsynaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (\seeFig{izhikevich}). We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task. 
%
\subsection{Application of the HD-SNN model to the motion detection task}
%
It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision. In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}. To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.

>>>>

To transform each movie into events, as recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames. For a given pixel and time stamp, an event is generated when the absolute value of this gradient exceeds a threshold. The event has either an OFF or ON polarity, depending on whether the gradient is negative or positive. The signed threshold is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, i.e. a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}). The goal here is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
%----------------------------%
%
\begin{figure}%[t!]
    \centering
    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons. Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}. When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function. In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}. The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron. An output spike is emitted when the membrane potential reaches the threshold and it is then reset.   \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses). }%
    \label{fig:izhikevich}%
\end{figure}%
% 
%----------------------------%
In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}. This can be formalized as a list of tuples, each tuple representing neural addresses and timestamps. We defined this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration). Events are usually ordered by their time of occurrence. Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$. This defines an address space $\presynaddrspace$, which consists of the set of possible addresses. In a neurobiological recording, this may be the identified set of all presynaptic neurons. For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels. Thus, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.

In the HD-SNN model, neurons $\postsynaddr \in \postsynaddrspace$ are connected to presynaptic afferent neurons from $\presynaddrspace$. In biology, a single cortical neuron typically has several thousand synapses. Each synapse can be defined by its synaptic weight and its delay, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays. Note that a neuron can contact another afferent neuron with different delays through different synaptic connections. By scanning all $\postsynaddr$ neurons, we define the full set of $\Nsyn$ synapses, 
as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$. This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of a neuron with synaptic weights and delays).

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. This results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$: 
%
\begin{linenomath*}
\begin{equation}\label{eq:stream_b}
\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
\end{equation}
\end{linenomath*}
%
In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons. In simulations, however, it should be properly reordered. The activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions (\seeFig{izhikevich}). Crucially, when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, the discharge probability increases, and especially when these spikes converge at the soma in a synchronous manner. We have defined the HD-SNN model in all generality in this subsection, and in the next subsection we describe an implementation of our model adapted to the motion detection task. 
%
\subsection{Application of the HD-SNN model to the motion detection task}
%
It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision. In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}. To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.
